---
title: "BU7152 Financial Analytics Indv Q2"
author: Matthew Forward
output: html_notebook
---

This notebook will look into the impacts of herding in two sectors: 

1. Canadian Banking

2. 7 Key NASDAQ Stocks

By looking into both, this notebook will be able to understand a potentially more
volatile stock(NASDAQ), that on the surface seems to have more potential for herding
than tradtional more stable stocks such as Canadian Bank returns.

Thus it will be interesting to see whether or not this initial hypothesis holds true.

## Loading Packages Required

Need to attach packages that will be required in running the subsequent code.
```{r}
library(quantmod)
library(TTR)
library (xts)
library (zoo)
library (PerformanceAnalytics)
library(pastecs)
library(dplyr)
library(Hmisc)
library (brms)
library(MSwM)
library(quantreg)

```


#Fetching Stock Data from YahooFinance

Acquiring 5 year data from YahooFinance on the Canadian Banking sector and key
NASDAQ stocks.

```{r}

#-----------Tech Stocks----------#
apple <- getSymbols("AAPL", src="yahoo", 
                         from=Sys.Date()-365*5,
                         to=Sys.Date(), auto.assign = FALSE)

facebook <- getSymbols("FB", src="yahoo", 
                         from=Sys.Date()-365*5,
                         to=Sys.Date(), auto.assign = FALSE)

alphabet <- getSymbols("GOOG", src="yahoo", 
                         from=Sys.Date()-365*5,
                         to=Sys.Date(), auto.assign = FALSE)

microsoft <- getSymbols("MSFT", src="yahoo", 
                     from=Sys.Date()-365*5,
                     to=Sys.Date(), auto.assign = FALSE)

tesla <- getSymbols("TSLA", src="yahoo", 
                     from=Sys.Date()-365*5,
                     to=Sys.Date(), auto.assign = FALSE)

amazon <- getSymbols("AMZN", src="yahoo", 
                     from=Sys.Date()-365*5,
                     to=Sys.Date(), auto.assign = FALSE)

nvidia <- getSymbols("NVDA", src="yahoo", 
                     from=Sys.Date()-365*5,
                     to=Sys.Date(), auto.assign = FALSE)

#---------Canadian Banks------------#

rbc <- getSymbols("RY", src="yahoo", 
                     from=Sys.Date()-365*5,
                     to=Sys.Date(), auto.assign = FALSE)
cibc <- getSymbols("CM", src="yahoo", 
                     from=Sys.Date()-365*5,
                     to=Sys.Date(), auto.assign = FALSE)
bmo <- getSymbols("BMO", src="yahoo", 
                     from=Sys.Date()-365*5,
                     to=Sys.Date(), auto.assign = FALSE)
td <- getSymbols("TD", src="yahoo", 
                     from=Sys.Date()-365*5,
                     to=Sys.Date(), auto.assign = FALSE)
scotiabank <- getSymbols("BNS", src="yahoo", 
                     from=Sys.Date()-365*5,
                     to=Sys.Date(), auto.assign = FALSE)
```


## Transform Closing Prices into 1 Dataframe

The only desired output from each time-series dataframe above is the closing price.
Closing price is extracted and the Canadian banks are put into one dataframe and NASDAQ
into the other.

```{r}
#--------Transform NASDAQ Stocks into One---------#

nasdaq <- cbind( apple$AAPL.Close, facebook$FB.Close, alphabet$GOOG.Close,
                 microsoft$MSFT.Close, tesla$TSLA.Close, amazon$AMZN.Close,
                 nvidia$NVDA.Close)
head(nasdaq)

#--------Transform NASDAQ Stocks into One---------#
canadaBanks <-cbind(rbc$RY.Close,cibc$CM.Close, bmo$BMO.Close,
                    td$TD.Close, scotiabank$BNS.Close)
summary(canadaBanks)
head(canadaBanks)
```



## Finding Log Returns

Here we want to find log daily returns across the sectors. Considering they're all 
taken from the same exchange the days will be the same (no issues with bank holidays
for different exchanges).

```{r}
#transform these to zoo
canadaBanks.zoo <- as.zoo(canadaBanks)
nasdaq.zoo <- as.zoo(nasdaq)


#---------find returns------#
returnNasdaq = Return.calculate( nasdaq , method = "log") 
returnCanadaBanks = Return.calculate( canadaBanks , method = "log") 


head(returnNasdaq)
head(returnCanadaBanks) #first row has NAs

summary(returnNasdaq)
summary(returnCanadaBanks)

```


## Interpreting Log Returns

Next we want to understand what the log returns are for each sector. The Canadian Banking
returns definintely have a tighter normal distribution with not as much ranging returns compared to
NASDAQ. This is expected, this does not necessarily imply that more herding occurrs in 
NASDAQ companies.

```{r}
#--- Understand the returns
canadaBanksDescriptive = stat.desc(returnCanadaBanks) # descriptive statistics of Canadina Banks
nasdaqDescriptive = stat.desc(returnNasdaq) # descriptive statistics of Nasdaq returns
nasdaqDescriptive
canadaBanksDescriptive


hist.data.frame(data.frame(returnNasdaq))
hist.data.frame(data.frame(returnCanadaBanks))
```



## Creating the CSAD Function and Calling It

Need to create this function to run create the market return and CSAD score for the groups
of equities (Banks and NASDAQ).

We then run the function to get CSAD and Rm. The means are lower for banking as expected
however there is the seems to not be much potential for herding for both as Rm^2 would
result in a lower number and may not be linearly dependent when we run regression.

Quick look at the plots and correlation for both with CSAD~Rm and CSAD~RM^2, it appears that
the strongest positive correlation exists between the Rm^2. Regeression will 
help determine whether the squared is significant.

```{r}
# a function to create CSAD and Rm
exchange.herd = function(return) 
{
  n=ncol(return)
  Rm = rowMeans(return)
  temp_dif =abs(return-Rm)
  temp_sum = rowSums(temp_dif)
  CSAD = temp_sum / ncol(return)
  CSAD = cbind (CSAD, Rm)
  return (CSAD)
}



nasdaqCSAD= exchange.herd(returnNasdaq)
banksCSAD = exchange.herd(returnCanadaBanks) # calling the function "exchange.herd" that calculates CSAD and Rm
head (banksCSAD) # show the first 6 rows
head(nasdaqCSAD)
summary(nasdaqCSAD)
summary(banksCSAD)

banksCSAD[,1]

#plot CSAD vs rm for canadina banks
plot(banksCSAD[,2], banksCSAD[,1])
plot(banksCSAD[,2]^2, banksCSAD[,1])
#correllations
cor(banksCSAD[-1,2], banksCSAD[-1,1])
cor(banksCSAD[-1,2]^2, banksCSAD[-1,1])

#plot CSAD vs rm NASDAQ
plot(nasdaqCSAD[,2], nasdaqCSAD[,1])
plot(nasdaqCSAD[,2]^2, nasdaqCSAD[,1])
#correlationss
cor(nasdaqCSAD[-1,2], nasdaqCSAD[-1,1])
cor(nasdaqCSAD[-1,2]^2, nasdaqCSAD[-1,1])

```




# Look for Herding in NASDAQ Funds

Here we want to determine whether normal regression is implying that NASDAQ stocks
have significant herding.

Evidently we see that Rm^2 is negative implying that some herding may be occurring,
however it is not statistically significant for us 
to determine that there is herding going on in the NASDAQ stocks. 

```{r}

#-----------find regular herding for NASDAQ-----------#
CSAD.df = fortify.zoo(nasdaqCSAD) # converting nasdaqCSAD into a dataframe (to simplify further calculations)
CSAD.df$Rm2 = CSAD.df$Rm^2 # calculating Rm^2
CSAD.df = CSAD.df[-c(1),] # removing the first row with NAs
head (CSAD.df) # show the first 6 rows
tail (CSAD.df) # show the last 6 rows

summary(CSAD.df)


y = CSAD.df$CSAD  # reassign my columns as Y and Xs to look better in the regression model
x1 = abs (CSAD.df$Rm)
x2 = CSAD.df$Rm2


#Linear model
linearMod <- lm(y~x1+x2)  # build linear regression model on full data
print(linearMod)
summary(linearMod)
```


## Fitting NASDAQ to other Regression Models

The Bayesian model indicates that RM^2 value will not be negative throughout
its entire 95% CI so its indicating that herding is not strong in NASDAQ stocks.

Similarly the quantile regression does not have a (-) RM^2 for 3 quantiles intervals,
and that ones that are negative are not statistically significant. 

Evidently we can generally say this basket of NASDAQ stocks does not show overall herding,
however we can still determine whether or not they follow (+) or negative herding.

```{r}
#bayesian
bayesian_data = cbind (y, x1, x2)
bayesian_model <- brm(formula = y ~ x1+x2,
                data = bayesian_data, seed = 99)
summary(bayesian_model)

#quantile regression
taus<- seq(from = .1, to = .9, by = .1)
coef0 <- rq( y ~ x1+x2, tau=taus);
summary(coef0);
plot (coef0, type = "l")

#markov switching model
nstates <-2 #setting this to two because of 1 clear irrational event:pandemic
msEuro = msmFit(linearMod , k = nstates , sw = rep(TRUE,4))
summary(msEuro)
msEuro
plotProb(msEuro, which =1)
```


## Find whether Herding Occurs Positively or Negatively in NASDAQ

First we need to create a dummy variable called dUP to determine whether or not
the overall market return was (+) on a day and label it 1 if it was and 0 if it wasnt.

Then we can introduce new variables to the regression to determine if there is (+) or 
(-) herding in group of NASDAQ securities.

Simple linear regression indicates that there is not statistically significant
evidence that (+) or (-) herding occurs. Although the values are negative (implying herding)
they are not significant to assume herding in either of the directions.

```{r}
#-------------finding positive/negative herding for NASDAQ---------#

CSAD.df = fortify.zoo(nasdaqCSAD) # converting f into a dataframe (to simplify further calculations)
CSAD.df$Rm2 = CSAD.df$Rm^2 # calculating Rm^2
CSAD.df <- CSAD.df %>%
  mutate(dUp=ifelse(Rm>0, 1, 0))
CSAD.df = CSAD.df[-c(1),] # removing the first row with NAs
head (CSAD.df) # show the first 6 rows
tail (CSAD.df) # show the last 6 rows

summary(CSAD.df)

# reassign  columns
y = CSAD.df$CSAD  # reassign  columns as Y and Xs 
x1 = abs(CSAD.df$Rm)*CSAD.df$dUp
x2 = abs(CSAD.df$Rm)*(1-CSAD.df$dUp)
x3 = CSAD.df$Rm2*CSAD.df$dUp
x4 = CSAD.df$Rm2*(1-CSAD.df$dUp)


head(CSAD.df)

#Linear model
linearMod <- lm(y~x1+x2+x3+x4)  # build linear regression model on full data
print(linearMod)
summary(linearMod)

```
## Running Quantile Regression on NASDAQ (+) (-) Herding

Quantile regression is run to indicate similar output to the linear regerssion, on 
determining whether (+) or (-) herding existed. It can be seen that most of the time,
X3 and X4 are not negative without signifcance thus assuring that the NASDAQ equities
do not behave (+) or (-) herding.

```{r}
#quantile regression
taus<- seq(from = .1, to = .9, by = .1)
coef0 <- rq( y ~ x1+x2+x3+x4, tau=taus);
summary(coef0);
plot (coef0, type = "l")

```














# Finding Herding in Canadian Banks

Next we want to determine whether herding occurs in Canadian Banks. From simple 
linear regression it can be determined that Canadian banks overall do not show herding
tendancies, as the RM^2 value is very small negative and is not statistically significant.

```{r}
#-----------find regular herding for Canadian Banks-----------#
CSAD.df = fortify.zoo(banksCSAD) # converting f into a dataframe (to simplify further calculations)
CSAD.df$Rm2 = CSAD.df$Rm^2 # calculating Rm^2
CSAD.df = CSAD.df[-c(1),] # removing the first row with NAs
head (CSAD.df) # show the first 6 rows
tail (CSAD.df) # show the last 6 rows

summary(CSAD.df)

# reassign my columns as Y and Xs to look better in the regression model
y = CSAD.df$CSAD  # reassign my columns as Y and Xs to look better in the regression model
x1 = abs (CSAD.df$Rm)
x2 = CSAD.df$Rm2


#Linear model
linearMod <- lm(y~x1+x2)  # build linear regression model on full data
print(linearMod)
summary(linearMod)


```
## Running other regression models on Canadian Banks Normal Herding

Similarly the coefficient  for the quantile method and bayesian regression models
indicate that Rm^2 is positive and when its (-) its statistically insignificant
implying that banks do not show herding tendancies, corroborating findings from normal
regression.

```{r}
#bayesian
bayesian_data = cbind (y, x1, x2)
bayesian_model <- brm(formula = y ~ x1+x2,
                data = bayesian_data, seed = 99)
summary(bayesian_model)

#quantile regression
taus<- seq(from = .1, to = .9, by = .1)
coef0 <- rq( y ~ x1+x2, tau=taus);
summary(coef0);
plot (coef0, type = "l")

```








## Determining whether or not there is positive or negative herding in Canadian Banks

The variables need to created and transformed similarly to what occurred for NASDAQ.
Once this happens a regression is run on these variables to try and indicate whether
there exists (+) or (-) herding.

From the results, x3 is positive and statistically significant, and x4 is not
statistically significant to imply any herding characteristics directionally.

Additionally we can see that the model still fits when looking for hederodeskacity,
so the lags and volatility of the errors is not affecting the regression model significantly.
The plot also shows the errors from the lm model do not follow any time series inference.

```{r}


#-------------finding positive/negative herding for Canadian Banks---------#
b = exchange.herd(banksCSAD)
head(b)
CSAD.df = fortify.zoo(b) # converting f into a dataframe (to simplify further calculations)
CSAD.df$Rm2 = CSAD.df$Rm^2 # calculating Rm^2
CSAD.df <- CSAD.df %>%
  mutate(dUp=ifelse(Rm>0, 1, 0))
CSAD.df = CSAD.df[-c(1),] # removing the first row with NAs
head (CSAD.df) # show the first 6 rows
tail (CSAD.df) # show the last 6 rows

summary(CSAD.df)

# reassign my columns as Y and Xs to look better in the regression model
y = CSAD.df$CSAD  # reassign my columns as Y and Xs to look better in the regression model
x1 = abs (CSAD.df$Rm)*CSAD.df$dUp
x2 = abs (CSAD.df$Rm)*(1-CSAD.df$dUp)
x3 = CSAD.df$Rm2*CSAD.df$dUp
x4 = CSAD.df$Rm2*(1-CSAD.df$dUp)


head(CSAD.df)

#Linear model
linearMod <- lm(y~x1+x2+x3+x4)  # build linear regression model on full data
print(linearMod)
summary(linearMod)

library(lmtest)
library(sandwich)
linearMod$coefficients
plot(linearMod$residuals)
coeftest(linearMod,vcov=NeweyWest(linearMod,verbose =T)) 



```


## Conclusion

It can be determined that the group of NASDAQ stocks and Canadian Banks do not show 
herding tendancies in any way.

NASDAQ has larger (-) coefficient for x2 for the regular herding test and
has larger (-) coefficient for x3,x4 for the (+) (-) herding test when comparing to
Canadian Banks. This was intiailly hypothesized, however due to the fact that these
coefficients are statistically insignifcant when performing regression analysis, it 
can be determined that the intial hypothesis was incorrect, and that both groups of
equities do not show herding tendancies.







