---
title: "Financial Analytics Factor Analysis"
author: "Matthew Forward"
output: html_notebook
---

# Looking at Facotr Analysis of Crypto Currencies and Traditional Financial Instruments

This notebook will be exploring the factors that relate between traditional financial instruments
and cryptocurrency. Correlations, economic factor models and statistical factor models will
be applied to a set of data.

### Loading Libraries Required

```{r}
#--------packages----------#
library(quantmod)
library(TTR)
library (xts)
library (zoo)
library (PerformanceAnalytics)
library(pastecs)
library(corrplot)
library (psych) # the library for factor analysis
library (GPArotation) # to estimate oblimin rotation

```


## Obtaining Financial Data

This section will be taking the last 5 years worth of data from YahooFinance, for
7 traditional financial instruments (dow jones index, nasdaq index, oil, gold,
us Treasury bonds, euro-USD and LSX index); additionally, 7 crpytocurrencies data is
collected (BTC,ETH,LTC, binance,cardano, monero, dogecoin).

The closing price data is then merged into one variable to perform analysis.

```{r}

#----------Downloading Data-----------#

#7 traditional
numberOfYears <-5
dow <- getSymbols("^DJI", src="yahoo", 
                    from=Sys.Date()-365*numberOfYears,
                    to=Sys.Date(), auto.assign = FALSE)

industrial <- getSymbols("^IXI", src="yahoo", 
                         from=Sys.Date()-365*numberOfYears,
                         to=Sys.Date(), auto.assign = FALSE)

usInterestRate <- getSymbols("^TNX", src="yahoo", 
                             from=Sys.Date()-365*numberOfYears,
                             to=Sys.Date(), auto.assign = FALSE)

nasdaq <- getSymbols("^IXIC", src="yahoo", 
                        from=Sys.Date()-365*numberOfYears,
                        to=Sys.Date(), auto.assign = FALSE)

oil <- getSymbols("BZ=F", src="yahoo", 
                    from=Sys.Date()-365*numberOfYears,
                    to=Sys.Date(), auto.assign = FALSE)

gold <- getSymbols("GC=f", src="yahoo", 
                  from=Sys.Date()-365*numberOfYears,
                  to=Sys.Date(), auto.assign = FALSE)



euroUSD <- getSymbols("EURUSD=X", src="yahoo", 
                             from=Sys.Date()-365*numberOfYears,
                             to=Sys.Date(), auto.assign = FALSE)
londonSX <- getSymbols("^FTSE", src="yahoo", 
                       from=Sys.Date()-365*numberOfYears,
                       to=Sys.Date(), auto.assign = FALSE)


tail(londonSX)

# 7 cryptos
BTC <- getSymbols("BTC-USD", src="yahoo", 
                       from=Sys.Date()-365*numberOfYears,
                       to=Sys.Date(), auto.assign = FALSE)

ETH <- getSymbols("ETH-USD", src="yahoo", 
                       from=Sys.Date()-365*numberOfYears,
                       to=Sys.Date(), auto.assign = FALSE)

LTC <- getSymbols("LTC-USD", src="yahoo", 
                     from=Sys.Date()-365*numberOfYears,
                     to=Sys.Date(), auto.assign = FALSE)

binance <- getSymbols("BNB-USD", src="yahoo", 
                     from=Sys.Date()-365*numberOfYears,
                     to=Sys.Date(), auto.assign = FALSE)

cardano <- getSymbols("ADA-USD", src="yahoo", 
                      from=Sys.Date()-365*numberOfYears,
                      to=Sys.Date(), auto.assign = FALSE)

monero <- getSymbols("XMR-USD", src="yahoo", 
                      from=Sys.Date()-365*numberOfYears,
                      to=Sys.Date(), auto.assign = FALSE)

dogecoin <- getSymbols("DOGE-USD", src="yahoo", 
                      from=Sys.Date()-365*numberOfYears,
                      to=Sys.Date(), auto.assign = FALSE)


# merge data and remove days where some have NAs

closingPrices <- cbind( dow$DJI.Close,industrial$IXI.Close, usInterestRate$TNX.Close,
                        nasdaq$IXIC.Close, oil$`BZ=F.Close`,
                        gold$`GC=F.Close`, euroUSD$`EURUSD=X.Close`,
                        londonSX$FTSE.Close, BTC$`BTC-USD.Close`, ETH$`ETH-USD.Close`,
                        LTC$`LTC-USD.Close`,binance$`BNB-USD.Close`,cardano$`ADA-USD.Close`,
                        monero$`XMR-USD.Close`,dogecoin$`DOGE-USD.Close`)

nrow(closingPrices) 
ncol(closingPrices)
```


## Transforming Data and Finding Log Returns
Considering their is missing data (mostly from commodities) and extra data (cryptocurrencies trade
365 days a year), some of the rows and data need to be omitted so then log returns can be performed 
on the same time series.

Log returns are then calculated on the closing price of days that have the same time frame.

In the 5 year dataset with 253 trading days/year, the dataset analyzed gets 1097/1265 possible
data points.

```{r}
#see that crpytos are traded over 365 days, while lots of missing days for traditional assets.
# need to omit entire rows of NAs before calculating logs.we get 1097/1265 - lots of issues
# with index and commodities.
summary(closingPrices)
nrow(closingPrices)

#omit entire row 
closingPrices <- na.omit(closingPrices)
nrow(closingPrices)
summary(closingPrices)
head(closingPrices)
tail(closingPrices)

#find log returns of omitted.
logReturn <- Return.calculate( closingPrices , method = "log") 
head(logReturn)
#get rid of first row of NAs
logReturn <- na.omit(logReturn)
nrow(logReturn)

```

## Correlation Testing

Next we want to look at correlation and determine what returns relate to eachother, this will give  a basic
understanding of what to expect from model factors.

Theres 5 observations:
1. The indexes all have a relatively strong positive correlation with one another and positive 
correlation with oil returns.
2. The cryptocurrencies mostly have a positve low correlation with all of the indexes.
3. The cyrptocurrencies have a strong positive correlation with one another.
4. Gold and US Treasury bonds are negatively correlated to one another and
have low  correlation with most other financial instruments.
5. Eur-USD currency exchange does not seem to be related to anything.

From these observations a rough guess of the amount of groups for factor analysis would be 
in the 3-5 range. More analysis will be done below.

```{r}
#corplot data to see which returns are strongly correlated with one another.
cor(logReturn)
corrplot(cor(logReturn), method = 'number')


```

# Factor Models:

## Economic Factor Model:

We will be looking at the economic factor model and the influence of the errors of three key
macroeconomic influences: US interest rates and industrial production index and dow jones index
and their affect on cryptocurrencies & traditional assets returns

Below we create a regression based off of the three influences mentioned and plot the r-squared
Overall the macro model does not fit that well to all of the variables. However on
an individual scale, the dow index, interest rates and industrial index show some resemblance
and factor into some of the returns of others.

Looking at pure rsquared, the LSX exchange and NASDAQ fit the best due to the fact that these are general
indexes so macro factors will have a higher relationshop to them compared to individiaul stocks
or cryptocurrency prices. This is expected when initially looking at correlation above.

Other observations that support above, the EURUSD and gold isnt really affected by
the effects of the dow whereas the other returns are more highly influenced.

Another observation is that most cryptos are influenced in same direction.

```{r}

## FACTOR Analysis
factors <- logReturn

macroFactors <- xts(logReturn[,1:8])

dow_arima <- arima(macroFactors[,1]) # autoregressive
dow_res <- dow_arima$residuals

interest_arima <- arima(macroFactors[,2]) # autoregressive
interest_res <- interest_arima$residuals

industrial_arima <- arima(macroFactors[,3]) # autoregressive
industrial_res <- industrial_arima$residuals

residuals <- cbind(industrial_res, dow_res, interest_res)


summary(residuals)

# residuals of the AR (5) modle (the interpretations of the residuals is that they are unexpected shocks)
lmfit = lm(factors[,4:15]~factors$IXI.Close+factors$DJI.Close+factors$TNX.Close) #fit a regression bitcoin is a set of teh dow, interest
# rate industrial and residuals from the AR models


slmfit = summary(lmfit) #summary of the regression
slmfit
rsq = rep(0,12) #create a variable rsq with seven 0 values


for (i in 1:12){rsq[i]= slmfit[[i]][[8]]} # we substruct the values of R2 from each of 7 models
beta_industrial = lmfit$coef[2,] # extract all industrial
beta_dow = lmfit$coef[3,] # extract all dow
beta_interest = lmfit$coef[4,] # extract all interest betas


#par(mfrow=c(1,12)) # building 4 graphs in a row
names(beta_dow)
barplot(rsq,horiz=T,names=names(beta_dow),main="R squared") #Creates a bar plot with vertical or horizontal bars.
barplot(beta_industrial,hori=T,main="beta Industrial") #Creates a bar plot with vertical or horizontal bars.
barplot(beta_dow,hori=T,main="beta Dow") #Creates a bar plot with vertical or horizontal bars.
barplot(beta_interest,hori=T,main="beta Interest") #Creates a bar plot with vertical or horizontal bars.
```







### Statistical Model: Determining if Data is Factorable 
First we want to determine whether or not our data is factorable.

First Bartletts Test of Sphericity is run to determine whether or not data is
factorable. Consdiering the P-value is 0, it means that data within intercorrelate with
one another, thus that means they could potentially be group together and factored.

Additionally KMO is run to determine  factorability of the overall model and the independent variables
Overall the model has a high MSA above the cutoff of 0.6 which means factor analysis can be run. One thing 
to note on an independent level, EUR-USD has a low MSA score, similar to our observation through 
correlations. We do not need to drop the variable at the moment, but aligns with initial obseravtion
that it may not be factorable or influence a factor in any significant manner.

```{r}
equities = logReturn
describe(equities) # general description of the data 

##Assessing the Factorability of the Data
#Bartlett's Test of Sphericity
cortest.bartlett(equities)

#KMO
KMO(equities)
```
## Determining Number of Factors to use in Model

Next a screen plot and parallel analysis scree plot is run to determine
how many factors should be used.

The scree plot indicates to use 4 factors, and parallel analysis 3-4.
Similar to our idea through the corrplots we will run factor analysis 
on 3,4 and maybe 5 factors and compare and see what is more interpretable
and understandable.




```{r}

# scree plot
scree(equities)

#Parallel Analysis
fa.parallel (equities) #



```



## Creating a 3 Factor Model

We want to create and analyze a 3 factor model.

The factors created are:
1. Cryptocurrency model
2. Industrial / Index Fund returns
3. Interest Rate and  negative Gold

Interestingly EURUSD and Oil are not included in the factors. These groupings are quite 
similar to initial observations made with analyzing the correlations.

The dow and index stocks communality are most explainable alongside the cryptocurrencies.

Subsequently the commodities dont have much communality thus the factors that are created in 3 factor 
model dont really explain commodites/fiat currency. Thus these factors probably shouldnt be used
if further analysis was to be used from in estimating currency/commodities.

Looking at uniqueness, it follows the communalities, in that gold, oil and EurUSD are
all high due to being unique and not having much multicollinearity. Whereas the dow has a
lot of multicolinearitys and the cryptos with eachtoher.

On interesting thing to note is that LSX index is relatively high for uniqueness(0.57)
ever though similar exchange like the dow is low.






```{r}
# estimation factor model
factor.model <- fa(equities, nfactors = 3, fm="ols", max.iter = 100, rotate = "oblimin")
# we estimate a factor model with 3 factors, estimated by means of OLS, 100 is the number of iterations or attempts to use when identifying the "best"" solution
# rotate - we apply oblimin rotation, allowing factors to correlate.
summary(factor.model)

# make it visual
fa.diagram(factor.model) # 

#loadings
factor.model$loadings
# Communality 
factor.model$communality

#Eeigenvalues
factor.model$e.values

#Percentage of Variance Accounted For
100*factor.model$e.values/length(factor.model$e.values)

#see how much infleunce that factors have
plot(1:15,100*factor.model$e.values/length(factor.model$e.values))

print(factor.model$loadings, cutoff=0, digits=3)
print(factor.model$Structure, cutoff=0, digits=3)
```


## Creating 4 Factor Model

We want to create and analyze a 4 factor model.

The 4 factor mnodel is highly similar to 3 factors with following groups

1. Cryptos
2. Dow, Nasdaq, and Industrial
3. London stock exchange, interest rate and brent crude
4. Gold

Similarly it leaves Euro/USD out of the model of significance. 

Factor 1 really explains  most cryptocurrencies well, besides dogecoin which is 
not as strong, this could be due to volatility and external influence of returns 
(ie Musk).

Factor 2 explains the index quite well, however its surpsiing that it does not explain
the LSX index well as one would expect that exchange to be similar to the US counterparts.

Factor 3 is just gold, and thus just putting into a model would be putting in price of
gold, does not help that much just that its of infleunce on its own.

Factor 4 is harder to interpret and think of reasons why.

Looking at the strctures, Factor 3 does not seem to do anything beyond gold and once again
eurusd is not really explained well. Thus the variance in EURUSD is not accounted for well.
Additionally we start seeing accounting of variance between factors suggesting they are
quite similar (Factors 2 & 4)


```{r}
# estimation factor model
factor.model <- fa(equities, nfactors = 4, fm="ols", max.iter = 100, rotate = "oblimin")

summary(factor.model)
# we estimate a factor model\

# make it visual
fa.diagram(factor.model) # 
#loadings
factor.model$loadings
# Communality 
factor.model$communality

#Eeigenvalues
factor.model$e.values

#Percentage of Variance Accounted For
100*factor.model$e.values/length(factor.model$e.values)

#see how much infleunce that factors have
plot(1:15,100*factor.model$e.values/length(factor.model$e.values))

print(factor.model$loadings, cutoff=0, digits=3)
print(factor.model$Structure, cutoff=0, digits=3)
```


## 5 Factors 
Quick look at 5 factor model, 5th factor does not have any values, thus we only need
to analyze a 3 and 4 factor model.




```{r}
# estimation factor model
factor.model <- fa(equities, nfactors = 5, fm="ols", max.iter = 100, rotate = "oblimin")

summary(factor.model)
# we estimate a factor model\

# make it visual
fa.diagram(factor.model) # 

#loadings
factor.model$loadings
```




# Conclusion


The 3 factor model seems to be the better of the factor models to use. The 5 factor model
produces a factor with no explainability to the variables thus this is disregarded.

The 3 factor makes more sense as it has indexes which are usually similar, gold and interest
rates which can be grouped together as something that is less not direct index returns, and 
cryptocurrencies grouped together, with EURUSD not as an influece.

The 4 factor model shouldnt be used as 1 factor explains only 1 variable, thus the factor
only tells about gold and having that additional factor does not make it overall more explainable.
Additionally in the 4 factor model there is more explainability between factor 2 and 4,
and thus these factors become quite related, which doesnt help when trying to differentiate 
different groups. Interstingly the dow had the highest eigenvalue and was thus the 
most explainable by the factors. Probably represenative as the dow is a barometer of macro
factors and performing stocks across most industries.

Thus the numeric recommends 3-4 models, and based off of understandability and what the models
create, a 3 model soultion is likely the best model.

the 3 factor model could be used to generally understand the markets in terms of general indexes
and cryptocurrency returns, one should be hesitant with both the 3&4 factor models to use it
for further investigations into fiat currencies, as the log returns on EURUSD was not influenced by
any of the factors created.

Overall the factor models were similar to initial observations when analzying correlation matrix, 
thus the factor model could be used in scenarios mentioned above.














